---
title: "Pcadapt workflow chr17"
author: "Luke Scott"
date: "`r Sys.Date()`"
output: pdf_document
---
* This is a workflow for identifying SNPs under selection using pcadapt.
See: https://bcm-uga.github.io/pcadapt/articles/pcadapt.html
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

* Load packages and read bed file
```{r}
library(pcadapt)
library(ggplot2)
library(qqman)
library(ggrepel)
library(dplyr)
test17 <- read.pcadapt("chr17.bed", type = "bed")
```

* Determine K value
```{r}
x <- pcadapt(input = test17, K = 5)
class(x)
summary(x)
plot(x, option = "screeplot") # K = 2
```

* Compute test statistics
```{r}
x1 <- pcadapt(test17, K = 2)
summary(x1)
plot(x1 , option = "manhattan")
plot(x1, option = "qqplot")
hist(x1$pvalues, xlab = "p-values", main = NULL, breaks = 50, col = "orange")
plot(x1, option = "stat.distribution")
```

* Check for linkage disequilibrium, indicated by clusters of SNPs.
```{r}
for (i in 1:2)
  plot(x1$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```


* Set method for p value adjustment and alpha value
```{r}
padjbonf <- p.adjust(x1$pvalues,method="bonferroni")
alpha <- 0.0001
outliersbonf <- which(padjbonf < alpha)
```

* Number of SNP outliers and SNP outlier list
```{r}
length(outliersbonf)
outliersbonf
```

* The association of SNPs with PCs
```{r}
get.pc(x1, outliersbonf)
```

* Plotting results with qqman
```{r}
summary(x1)
#make dataframe with values from pcadapt
SNP <- x1$pass
CHR <- rep(17, times = 9093) 
BP <- x1$pass
P <- x1$pvalues[!is.na(x1$pvalues)]

qqdf17 <- data.frame(SNP, CHR, BP, P)

manhattan(qqdf17, main = " pcadapt SNP Outliers", cex.axis = 0.8, cex.main = .8, 
          annotatePval = 0.0000001, suggestiveline = F , annotateTop = FALSE, 
          xlab = "Chromosome 17", cex = 0.3, highlight = outliersbonf)

```

* Plotting results with ggplot
```{r}
# Prepare the dataset
don <- qqdf17 %>% 
  
  # Compute chromosome size
  group_by(CHR) %>% 
  summarise(chr_len=max(BP)) %>% 
  
  # Calculate cumulative position of each chromosome
  mutate(tot=cumsum(chr_len)-chr_len) %>%
  select(-chr_len) %>%
  
  # Add this info to the initial dataset
  left_join(qqdf17, ., by=c("CHR"="CHR")) %>%
  
  # Add a cumulative position of each SNP
  arrange(CHR, BP) %>%
  mutate( BPcum=BP+tot) %>%
  
  # Add highlight and annotation information
  mutate( is_highlight=ifelse(SNP %in% outliersbonf, "yes", "no")) %>%
  mutate( is_annotate=ifelse(-log10(P)>8, "yes", "no")) 

# Prepare X axis
axisdf <- don %>% group_by(CHR) %>% summarize(center=( max(BPcum) + min(BPcum) ) / 2 )

# Make the plot
ggplot(don, aes(x=BPcum, y=-log10(P))) +
  
  # Show all points
  geom_point( aes(color=as.factor(CHR)), alpha=0.8, size=1.3) +
  scale_color_manual(values = rep(c("black"))) +
  
  # custom X axis:
  scale_x_continuous( label = axisdf$CHR, breaks= axisdf$center ) +
  scale_y_continuous(expand = c(0, 0) ) +     # remove space between plot area and x axis
  
  # Add highlighted points
  geom_point(data=subset(don, is_highlight=="yes"), color="orange", size=2) +
  
  # Add label using ggrepel to avoid overlapping
  geom_label_repel( data=subset(don, is_annotate=="yes"), aes(label=SNP), 
                    size=2, max.overlaps = 15) +
  #horizontal line
  geom_hline(yintercept = 8, col="red", alpha = 0.5) +
  
  labs(title = "pcadapt SNP outliers", 
       x = "Chromosome 17 SNPs") +
  
  # Customize theme:
  theme_bw() +
  theme( 
    legend.position="none",
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

```

